---
title: "Human Development Analysis"
subtitle: Jenn Griffiths Data Science Portfolio
output:
  html_document:
    df_print: paged
---
#Install Libraries
These libraries allow us to use the various functions within this document
```{r echo=TRUE, message=FALSE, warning=TRUE, paged.print=FALSE}
include <- function(library_name){
  if( !(library_name %in% installed.packages()) )
    install.packages(library_name) 
  library(library_name, character.only=TRUE)
}
library(data.table)
#janitor package for cleaning column names
include("janitor")
# having errors when trying to load tidyverse
#include("tidyverse")
include("caret")
include("knitr")
include("readr")
include("dplyr")
include("tidyr")
include("ggplot2")
```
#Load the Data
Loading the original data into the analysis environment. The original data will not be altered. Versions of the data are created in this environment which can be freely manipulated. If an alteration, such as a removal of a column, was a mistake - one can always reload the original data. 
```{r include=FALSE}
Indicators <- read_csv("Data/Indicators.csv")
#An additional Dataset in the form of a CSV
HDI <- read_csv("Data/HDI.csv")
#Lets remove the first column , as it is a useless duplicate index
HDI[1] <- NULL
```
#View the Dataframes
This function will show us the top 6 rows of the data frame in order to give us an idea of the structure of table. 
```{r}
head(Indicators)
```
#Inspecting the Data
```{r}
#Count of rows
nrow(Indicators)
#Count of columns
ncol(Indicators)
```
###List of the variables in the dataframe
```{r}
names(Indicators)
```
#Tidy The Data
Tidying the data is an important part of preparing your data for analysis. Making your data tidy can be time consuming but it makes the format of your data consistent and easier to manage when you're analyzing the data.

###Categorical Variables
Change variable 'dimension' to a categorical variable, meaning there are only a certain number of values it could possibly be. This will help with analysis later. 

```{r}
Indicators$dimension <-as.factor(Indicators$dimension)

#Display number of levels in 'dimension' variable
nlevels(Indicators$dimension)
```
>Change variable 'indicator_name' to a categorical variable

```{r}
Indicators$indicator_name <-as.factor(Indicators$indicator_name)

#Display levels in 'indicator_name' variable
nlevels(Indicators$indicator_name)
```
>Change variable 'country_name' to a categorical variable

```{r}
Indicators$country_name <-as.factor(Indicators$country_name)

#Display levels in 'country_name' variable
nlevels(Indicators$country_name)
```
>Change variable 'indicator_id' to a categorical variable

```{r}
Indicators$indicator_id <-as.factor(Indicators$indicator_id)
```
>Change variable 'iso3' to a categorical variable

```{r}
Indicators$iso3 <-as.factor(Indicators$iso3)
```
###Gather year columns
An important feature of tidy data is that "each variable must have its own column" As in the variable years should be in a single column. This is done because it makes transforming the data a much smoother process. 
```{r}
Indicators <-gather(Indicators, key = Years, "HDI_value", 6:34, na.rm = TRUE)

#rename last column to rating
#colnames(Indicators)[colnames(Indicators) == "1990" ] <- "HDI_value"
```
#Subset the data based on 'dimensions' category
To break up the data by 'dimensions', we must take subsets of the original table. It is good practice to have each table be one thing, rather than a collection of everything. Dimensions in this data set are the category under which the Human Development Index is scored. For example dimensions include education, health, demographics, and so on. 

```{r}
Composite_indices <- subset(Indicators, dimension == "Composite indices" )
Composite_indices <- as_tibble(Composite_indices, na.rm = TRUE)

Demography <- subset(Indicators, dimension == "Demography" )
Demography <- as_tibble(Demography, na.rm = TRUE)

Education <- subset(Indicators, dimension == "Education" )
Education <- as_tibble(Education, na.rm = TRUE)

Environmental_sustainability <- subset(Indicators, dimension == "Environmental sustainability" )
Environmental_sustainability <- as_tibble(Environmental_sustainability)

Gender <- subset(Indicators, dimension == "Gender" )
Gender <- as_tibble(Gender, na.rm = TRUE)

Health <- subset(Indicators, dimension == "Health" )
Health <- as_tibble(Health, na.rm = TRUE)

Human_Security <- subset(Indicators, dimension == "Human Security" )
Human_Security <- as_tibble(Human_Security, na.rm = TRUE)

Income_resources <- subset(Indicators, dimension == "Income/composition of resources" )
Income_resources <- as_tibble(Income_resources, na.rm = TRUE)

Inequality <- subset(Indicators, dimension == "Inequality" )
Inequality <- as_tibble(Inequality, na.rm = TRUE)

Mobility_communication <- subset(Indicators, dimension == "Mobility and communication" )
Mobility_communication <- as_tibble(Mobility_communication,na.rm = TRUE)

Poverty <- subset(Indicators, dimension == "Poverty" )
Poverty <- as_tibble(Poverty, na.rm = TRUE)

Sustainability <- subset(Indicators, dimension == "Socio-economic sustainability" )
Sustainability <- as_tibble(Sustainability, na.rm = TRUE)

Financial_flows <- subset(Indicators, dimension == "Trade and financial flows" )
Financial_flows <- as_tibble(Financial_flows, na.rm = TRUE)

Work_vulnerability <- subset(Indicators, dimension == "Work, employment and vulnerability" )
Work_vulnerability <- as_tibble(Work_vulnerability, na.rm = TRUE)
```


#Now It's Time For STEP ZERO: Plot The Data
Nothing gives you a better perspective on your data than plotting your data. 

###Bar Chart across 2 s
At first glance the two tall bars in the center look as though there is something weird going on here. In looking into the tibble, it is evident that the HDI_value is scoring each indicator with different units. Most scores are in context of percentages, whereas the two tall bars are in terms of headcounts of certain population groups. This leads to the conclusion that these tibbles need to be sliced again based on the indicators in order for the HDI_values in a tibble to be in the same units.

> This plot demonstrates that the HDI value is not a uniform unit of measure. For each dimension (e.g. Education, Poverty, Health ) there are multiple sebcategories, Indicators that use different units of measure. 

```{r}
plot <- ggplot(Poverty, aes(indicator_name, HDI_value) ) + geom_col() + geom_path(colour = "red", na.rm = TRUE)
# Rename the x, Y axis to make graph communnicate more clearly
plot + ggtitle("Plot to demonstrate the HDI score is not a uniform unit of measure") + theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + xlab("Indicators Within A Dimension") + ylab("HDI Score")
```

###Bar Chart
Looking at distributions of a single variable
The assumption would be that there is an even spread of the amount of indicators across the data set but here we see there a subtle dip then a dramatic rise at the far right of the bar graph. 

Is this due to the missing data being removed? Or possibly some other reason for the uneven distribution?

```{r}

chart<- ggplot(Poverty, aes(indicator_name), na.rm = TRUE)
chart + geom_bar() + theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())+ ggtitle('Indicators Have a Slight Variation In Occurrences')
```

This graph looks at the HDI scores in terms of each country:
```{r}
ggplot(Poverty , aes(x = country_name, y = HDI_value)) + geom_col() + theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())+ ggtitle("HDI Scores Among Each Country")
```

