---
title: "Human Development Analysis"
subtitle: "Jenn Griffiths's Data Science Portfolio" 
output: html_notebook
---
#Install Libraries
```{r}
#install.packages("tidyverse")
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
```
#Load the Data
```{r}
HDI <- read_csv("Data/HDI.csv")
Continent_HDI <- read_csv("Data/Continent_HDI.csv")
Indicators <- read_csv("Data/Indicators.csv")
```
#View the Dataframes
```{r}
View(Continent_HDI)
View(HDI)
View(Indicators)
```
#Inspecting the Data
```{r}
#Count of rows
nrow(Indicators)
#Count of columns
ncol(Indicators)
#List the variables in dataframe
names(Indicators)
```
>Summary Statistics

```{r}
summary(Indicators)
```
>List the structure of dataframe

```{r}
glimpse(Indicators)
```
#Tidy data for Indicators dataframe
>Change variable 'dimension' to a factor

```{r}
Indicators$dimension <-as.factor(Indicators$dimension)

#Display levels in 'dimension' variable
levels(Indicators$dimension)
```
>Change variable 'indicator_name' to a factor

```{r}
Indicators$indicator_name <-as.factor(Indicators$indicator_name)

#Display levels in 'indicator_name' variable
levels(Indicators$indicator_name)
```
>Change variable 'country_name' to a factor

```{r}
Indicators$country_name <-as.factor(Indicators$country_name)

#Display levels in 'country_name' variable
levels(Indicators$country_name)
```
>Change variable 'indicator_id' to a factor

```{r}
Indicators$indicator_id <-as.factor(Indicators$indicator_id)

#Display levels in 'indicator_id' variable
levels(Indicators$indicator_id)
```
>Change variable 'iso3' to a factor

```{r}
Indicators$iso3 <-as.factor(Indicators$iso3)

#Display levels in 'iso3' variable
levels(Indicators$iso3)
```
>Verify the structure of dataframe

```{r}
glimpse(Indicators)
```

#Adressing large amount of 'missing' data
```{r}
#is.na(Indicators)
Indicators[!complete.cases(Indicators),]

#create new dataset without missing data
Indicator_noNa<-na.omit(Indicators)
glimpse(Indicator_noNa)
```
